---
title: "Week 1"
output: html_notebook
---

## Exploratory data analysis and visualization 

Loading the data
```{r}
orgs <- read.csv("organisms_genes.tsv", sep='\t')
```

Preview the data
```{r}
head(orgs)
table(orgs$type)
```

maybe plot out the number of genes that code for proteins over the total number of basepairs
```{r}
library(ggplot2)
ggplot(orgs, aes(bp, coding, color=type)) +
      geom_point(size=3)
```
there's one organisms has significantly higher number of both basepairs and genes that code for proteins, but what is it?
```{r}
orgs[which.max(orgs$bp),]
mean(orgs$bp)
```

to make the plot more interpretable, you can just remove it
```{r}
orgs_filter <- orgs[orgs$name != "Wheat",]
ggplot(orgs_filter, aes(bp, coding, color=type)) +
        geom_point(size=3)
```

or taking the log
```{r}
orgs$log_bp <- log10(orgs$bp)
orgs$log_coding <- log10(orgs$coding)
library(dplyr)
orgs <- orgs %>% mutate(log_bp = log10(bp),
                        log_coding = log10(coding))
head(orgs)
ggplot(orgs, aes(log_bp, log_coding, color=type)) +
  geom_point(size=3)
```

let's try to make a new variable and calculate the $mu$ and $sigma$ of the ratio of genes which are coding 
```{r}
orgs <- orgs %>% mutate(genes = coding + noncoding)
orgs %>%
  group_by(type) %>%
  summarize(n=n(),
            coding_ratio_mu = mean(coding/genes),
            coding_ratio_sigma = sd(coding/genes))
```
and have some more plots to visulize the this data, like the range of bp for each organism type
```{r}
orgs_filter %>%
  group_by(type) %>%
  summarize(basepairs=median(bp),
            min=min(bp),
            max=max(bp)) %>%
  ggplot(aes(type, basepairs, ymin=min, ymax=max)) +
          geom_pointrange()
```
or do some organism-type-wise linear regression fitting on the bp vs coding plots
```{r}
ggplot(orgs_filter, aes(bp, coding, col=type)) +
        geom_point() + geom_smooth(se=FALSE, method="lm")
```
or have “faceting” plots by types
```{r}
ggplot(orgs, aes(bp, coding)) +
  geom_point() + facet_wrap(~ type, scales="free")
```
