---
title: "Week 2"
output: html_notebook
---

First, install all the necessary packages
```{r}
BiocManager::install("GEOquery")
BiocManager::install("preprocessCore")
install.packages('pheatmap')
install.packages('matrixStats')
```
and attach them
```{r}
library(GEOquery)
library(preprocessCore)
library(matrixStats)
library(pheatmap)
```

Download the expression data using `GEOquery`
```{r}
file <- list.files(pattern="GSE2125")
if (length(file) == 0) {
  geo <- getGEO("GSE2125", destdir=".")
  e <- geo[[1]]
} else {
  e <- getGEO(filename=file)
}
```
Let's have a look at this data we just downloaded
```{r}
head(exprs(e))
```
```{r}
head(pData(e))
```
```{r}
head(fData(e))
```
```{r}
e$extract_protocol_ch1[1]
```
```{r}
e$characteristics_ch1
```
```{r}
table(e$characteristics_ch1)
```

check what's the range of the gene expression and draw a boxplot
```{r}
range(exprs(e))
boxplot(exprs(e),range=0)
```
we see there're some difference in expression ranges between different gene, better to preprocess it with `normalize.quantiles`
```{r}
exprs(e) <- normalize.quantiles(exprs(e))
boxplot(exprs(e),range=0)
```
Finally, let's make a heatmap of the expression data
```{r}
exp_val <- rowVars(exprs(e))
o <- head(order(exp_val, decreasing=TRUE),200)
pheatmap(exprs(e)[o,],
         annotation_col=pData(e)["characteristics_ch1"],
         show_rownames=FALSE, show_colnames=FALSE)
```

