---
title: "Week 3"
output: html_notebook
---
```{r}
# loading packages
# BiocManager::install(c("Biostrings", "BSgenome.Hsapiens.NCBI.GRCh38"))
library(Biostrings)
library(BSgenome.Hsapiens.NCBI.GRCh38)
```

### DNAString object
It encodes a DNA string. You can measure its length with `length()` or get its reverse complement by `reverseComplement`.
```{r}
dna_str <- DNAString("AAACGCG")
dna_str
length(dna_str)
reverseComplement(dna_str)
```

### DNAStringSet object
It encodes a collection of DNA strings. You can measure the number of DNA strings with `length()` and the length of each string with `width()`.
```{r}
dna_set <- DNAStringSet(c("AAAAA","CGCG","TCG"))
dna_set
reverseComplement(dna_set)
length(dna_set)
width(dna_set)
```

### Useful functions
1. occurence/frequency
`letterFrequency()` measures the occurence of a signle letter or a pattern in a dna string
```{r}
letterFrequency(dna_str, 'C')
letterFrequency(dna_set, 'C')
letterFrequency(dna_set, "C", as.prob=TRUE)
letterFrequency(dna_str, "CG")
letterFrequency(dna_set, "CG")
```
2. matching
```{r}
# single pattern + multiple strings
dna <- DNAStringSet(c("AACTCTC","CTCTAAA","AAAGAG"))
matches <- vmatchPattern("CTC", dna)
matches
elementNROWS(matches)
# multiple pattern + single string
dna <- DNAString("AAACTCAAAGAGAAATTTAAA")
pd <- PDict(c("CTC","GAG","TTT","AAA"))
matches <- matchPDict(pd, dna)
matches
elementNROWS(matches)
```

### Example
Full genome sequences for Homo sapiens
```{r}
Hsapiens
seqnames(Hsapiens)
seqinfo(Hsapiens)
```
play with chromosome 1
```{r}
chr1 <- Hsapiens[[1]]
length(chr1)
alphabetFrequency(chr1)
matchPattern("CTC", chr1)
```
match pattern allowing mismatches
```{r}
n_pattern = "AAGCCTAAGCCTAAGCCTAA"
matchPattern(seq_pattern, chr1)
matches <- matchPattern(n_pattern, chr1, max.mismatch=2)
matches
mismatch(n_pattern, matches)
```
select sequences at given location
```{r}
gr <- GRanges("1", IRanges(1e6 + c(1,101,201), width=100))
getSeq(Hsapiens, gr)
```
