---
title: "Week 5"
output: html_notebook
---

```{r}
# BiocManager::install("RTCGA.clinical")
# BiocManager::install("RTCGA.mRNA")
library(RTCGA.clinical)
library(RTCGA.mRNA)
library(tibble)
```

### Survival Analysis
#### survival analysis using only 1 feature
```{r}
clin <- survivalTCGA(BRCA.clinical, OV.clinical, GBM.clinical, 
                     extract.cols="admin.disease_code")
head(clin)
```

get the count for each disease and contingency tabel
```{r}
table(clin$admin.disease_code)
xtabs(~admin.disease_code+patient.vital_status, data=clin)
```
fit Cox model and generate Kaplan–Meier plot
```{r}
# cox
coxph(Surv(times, patient.vital_status)~admin.disease_code, data=clin)
# km
sfit <- survfit(Surv(times, patient.vital_status)~admin.disease_code, data=clin)
summary(sfit, times=seq(0,365*5,365))
ggsurvplot(sfit, conf.int=TRUE, pval=TRUE)
```

#### survival analysis using only multiple feature
```{r}
clinkid <- survivalTCGA(KIPAN.clinical, 
                     extract.cols=c("admin.disease_code", "patient.gender"))
head(clinkid)
```
get the contingency table
```{r}
xtabs(~admin.disease_code+patient.gender, data=clinkid)
```
fit Cox model and generate Kaplan–Meier plot
```{r}
# cox
coxph(Surv(times, patient.vital_status)~admin.disease_code+patient.gender, data=clinkid)
# km
sfit <- survfit(Surv(times, patient.vital_status)~admin.disease_code+patient.gender, data=clinkid)
summary(sfit, times=seq(0,365*5,365))
ggsurvplot(sfit, conf.int=TRUE, pval=TRUE)
```
### more clustering example: kmenas using TCGA data
select the top 200 most variable genes
```{r}
data <- rbind(BRCA.mRNA, OV.mRNA)[, 2:17815]
std <- apply(data, 2, sd, na.rm =TRUE)
top200 <- order(std, decreasing = TRUE)[1:200]
gene_names <- colnames(data)
selected_genes <- gene_names[top200]
expr <- expressionsTCGA(BRCA.mRNA, OV.mRNA,
                        extract.cols = selected_genes)
```
fitting kmeans
```{r}
expr <- na.omit(expr)
km.res <- kmeans(expr[3:202], centers = 2)
str(km.res)
table(true=expr$dataset,cluster=km.res$cluster)
```



